Increasing the topological gap by an order of magnitude through the geometry for Majorana SNS junctions.

### 2019-02-07 paper/snakemajoranas.tex
align Hamiltonian eq
comment out the analytics and improve the abstract

\date{\today}
\begin{abstract}
We solve the problem of soft gap in high density SNS Majorana junctions \cite{pientka2017topological}.
Trajectories that do not encounter the superconductor for a long time weaken proximity effect in Josephson junctions.
A small induced gap becomes then an important obstacle in creating Majorana states.
We show that a zigzag-shaped geometry eliminates these trajectories, at the same time allowing robust creation of Majorana states with a gap increased by more than an order of magnitude.
\end{abstract}

@@

\begin{small}
\begin{align}
    H = \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right)\tau_z+
        E_\text{z} \sigma_x+
        \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z +
        \Delta \tau_y
\end{align}
\label{eq:hamiltonian}
\end{small}

@@

%%██████████████████████████████████████████████████████████████████████████
%%██ Analytical estimate
%%██████████████████████████████████████████████████████████████████████████
\section{Analytical estimate of the band gap}

\comment{To compute the effect of cutoff consider a single segment of the zigzag and obtain the spectrum analytically (this allows to use translation invariance).}
We establish that the cutoff of long trajectories is the driving mechanism behind the order of magnitude gap improvement using an analytical approach.
We derive the Andreev spectrum for a single (diagonal) segment of the zigzag, and define the gap of the system as the minimum of the spectrum occurring before the cutoff corresponding to the particular geometry.
The system under consideration is thus a ribbon SNS junction with a tilted magnetic field.

\comment{The definition of the cutoff...}

\comment{We use a short junction approximation and neglect transverse SO.}
We derive the spectrum by calculating the scattering matrix for the normal region and applying the Andreev bound state condition for the short junction limit\cite{beenakker1991universal, sticlet_robustness_2017}.
For derivation of the scattering matrix, we neglect SOI in the transverse direction, which is valid for $W<l_\text{so}$.
The result is specified in Eqs.~\eqref{eq:smatrix} to~\eqref{eq:momenta_short_junction}.

\begin{align}
\label{eq:smatrix}
    S &= \left(
    \begin{array}{rr}
    r_{ll}&t_{rl}\\
    t_{lr}&r_{rr}\\
    \end{array}
    \right) =
    \left(
    \begin{array}{rr}
    \beta_+ & e^{-i q W} \beta_-\\
    e^{-i q W} \beta_- & e^{-2 i q W} \beta_+\\
    \end{array}
    \right)
    \\
    \beta_\pm &= \left(
    \begin{array}{rr}
    e^{i \nu_{\arg}}\left(\omega^\pm_1 - \omega^\pm_2\right) & (\omega^\pm _1 + \omega^\pm _2)\\
    -(\omega^\pm _1 + \omega^\pm _2) & e^{-i \nu _{\arg }} \left(\omega^\pm _2 - \omega^\pm _1\right)\\
    \end{array}
    \right)\\
    \omega^\pm_j &= \frac{1}{4} \left(\gamma _{j} \pm \delta _{j}\right)
\end{align}

\begin{align}
    \gamma_j &= \frac{q+i k_{j} \tan \left(\frac{W k_{j}}{2}\right)}{q-i k_{j} \tan \left(\frac{W k_{j}}{2}\right)} \\
    \delta_j &= \frac{q-i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}{q+i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}\\
    e^{i \nu_{\arg}} &= \frac{E_\text{z} e^{i \theta }-i \alpha  k_x}{\sqrt{E_\text{z}^2+\alpha  k_x \left(\alpha  k_x-2 E_\text{z} \sin (\theta )\right)}}
\end{align}

\begin{footnotesize}
\begin{align}
    q &= \left[ \frac{2 m_\ast}{\hbar ^2}\mu_s - k_x^2 \right]^\frac{1}{2}\\
    k_1 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n-\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}\\
    k_2 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n+\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}
\label{eq:momenta_short_junction}
\end{align}
\end{footnotesize}

In order to get the spectrum from the S-matrix [Eq.~\eqref{eq:smatrix}] we use the main result from Ref.~\onlinecite{beenakker1991universal} and start with a determinantal equation for the bound state energies in a SNS-junction:

\begin{equation}
\det\left[1+\alpha^{2}\left(E\right)r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)\right]=0
\end{equation}

We rewrite this as a characteristic polynomial problem $\det\left[A-\lambda I\right]=0$, as

\begin{equation}
\det\left[r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)-\frac{-1}{\alpha^{2}\left(E\right)}I\right]=0
\end{equation}

with
\begin{equation}
\lambda=-\frac{1}{\alpha^{2}\left(E\right)},\;A=r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right).
\end{equation}

Where $\lambda_i$ are the eigenvalues of $A$.
Inverting $\alpha\left(E\right)\equiv\exp\left(-i\arccos\left(E/\Delta\right)\right)$ yields $\frac{E}{\Delta}=\frac{\alpha^{2}+1}{2\alpha}=\frac{1}{2}\left(\alpha+\alpha^{-1}\right)=\textrm{Re}(\alpha)$, where the last equality holds because $\alpha$ is unitary.
Then, since $\alpha$ is only defined in the positive imaginary plane, the energies are $\frac{E_{i}}{\Delta}=\textrm{Re}\left(\sqrt{-1/\lambda_{i}}\right)$ where we take the root which has a positive imaginary part.
Using this result and Eq.~\eqref{eq:smatrix}, we numerically find the Andreev spectrum.

\comment{The lowest energy at $|k| < k_\textrm{cutoff}$ sets the band gap.}
We cut off the spectrum at the momentum corresponding to the proposed cutoff in the zigzag, and compare it to the numerically obtained gap for a true zigzag system.

>>>>

\date{\today}
\begin{abstract}
We solve the problem of the appearance of a soft gap in high-density SNS Majorana junctions \cite{pientka2017topological}.
Trajectories that do not encounter the superconductor for a long time weaken the proximity effect in Josephson junctions.
A small induced gap then becomes an essential obstacle in creating Majorana states.
We show that a zigzag-shaped geometry eliminates these trajectories, at the same time allowing the robust creation of Majorana states with a gap increased by more than an order of magnitude.
\end{abstract}

@@

\begin{equation}
\begin{split}
    H = & \left(\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right)\tau_z \\
        & + E_\text{z} \sigma_x
          + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z
          + \Delta \tau_y
\end{split}
\label{eq:hamiltonian}
\end{equation}

@@

%%██████████████████████████████████████████████████████████████████████████
%%██ Analytical estimate
%%██████████████████████████████████████████████████████████████████████████
% \section{Analytical estimate of the band gap}

% \comment{To compute the effect of cutoff consider a single segment of the zigzag and obtain the spectrum analytically (this allows to use translation invariance).}
% We establish that the cutoff of long trajectories is the driving mechanism behind the order of magnitude gap improvement using an analytical approach.
% We derive the Andreev spectrum for a single (diagonal) segment of the zigzag, and define the gap of the system as the minimum of the spectrum occurring before the cutoff corresponding to the particular geometry.
% The system under consideration is thus a ribbon SNS junction with a tilted magnetic field.

% \comment{The definition of the cutoff...}

% \comment{We use a short junction approximation and neglect transverse SO.}
% We derive the spectrum by calculating the scattering matrix for the normal region and applying the Andreev bound state condition for the short junction limit\cite{beenakker1991universal, sticlet_robustness_2017}.
% For derivation of the scattering matrix, we neglect SOI in the transverse direction, which is valid for $W<l_\text{so}$.
% The result is specified in Eqs.~\eqref{eq:smatrix} to~\eqref{eq:momenta_short_junction}.

% \begin{align}
% \label{eq:smatrix}
%     S &= \left(
%     \begin{array}{rr}
%     r_{ll}&t_{rl}\\
%     t_{lr}&r_{rr}\\
%     \end{array}
%     \right) =
%     \left(
%     \begin{array}{rr}
%     \beta_+ & e^{-i q W} \beta_-\\
%     e^{-i q W} \beta_- & e^{-2 i q W} \beta_+\\
%     \end{array}
%     \right)
%     \\
%     \beta_\pm &= \left(
%     \begin{array}{rr}
%     e^{i \nu_{\arg}}\left(\omega^\pm_1 - \omega^\pm_2\right) & (\omega^\pm _1 + \omega^\pm _2)\\
%     -(\omega^\pm _1 + \omega^\pm _2) & e^{-i \nu _{\arg }} \left(\omega^\pm _2 - \omega^\pm _1\right)\\
%     \end{array}
%     \right)\\
%     \omega^\pm_j &= \frac{1}{4} \left(\gamma _{j} \pm \delta _{j}\right)
% \end{align}

% \begin{align}
%     \gamma_j &= \frac{q+i k_{j} \tan \left(\frac{W k_{j}}{2}\right)}{q-i k_{j} \tan \left(\frac{W k_{j}}{2}\right)} \\
%     \delta_j &= \frac{q-i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}{q+i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}\\
%     e^{i \nu_{\arg}} &= \frac{E_\text{z} e^{i \theta }-i \alpha  k_x}{\sqrt{E_\text{z}^2+\alpha  k_x \left(\alpha  k_x-2 E_\text{z} \sin (\theta )\right)}}
% \end{align}

% \begin{footnotesize}
% \begin{align}
%     q &= \left[ \frac{2 m_\ast}{\hbar ^2}\mu_s - k_x^2 \right]^\frac{1}{2}\\
%     k_1 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n-\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}\\
%     k_2 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n+\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}
% \label{eq:momenta_short_junction}
% \end{align}
% \end{footnotesize}

% In order to get the spectrum from the S-matrix [Eq.~\eqref{eq:smatrix}] we use the main result from Ref.~\onlinecite{beenakker1991universal} and start with a determinantal equation for the bound state energies in a SNS-junction:

% \begin{equation}
% \det\left[1+\alpha^{2}\left(E\right)r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)\right]=0
% \end{equation}

% We rewrite this as a characteristic polynomial problem $\det\left[A-\lambda I\right]=0$, as

% \begin{equation}
% \det\left[r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)-\frac{-1}{\alpha^{2}\left(E\right)}I\right]=0
% \end{equation}

% with
% \begin{equation}
% \lambda=-\frac{1}{\alpha^{2}\left(E\right)},\;A=r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right).
% \end{equation}

% Where $\lambda_i$ are the eigenvalues of $A$.
% Inverting $\alpha\left(E\right)\equiv\exp\left(-i\arccos\left(E/\Delta\right)\right)$ yields $\frac{E}{\Delta}=\frac{\alpha^{2}+1}{2\alpha}=\frac{1}{2}\left(\alpha+\alpha^{-1}\right)=\textrm{Re}(\alpha)$, where the last equality holds because $\alpha$ is unitary.
% Then, since $\alpha$ is only defined in the positive imaginary plane, the energies are $\frac{E_{i}}{\Delta}=\textrm{Re}\left(\sqrt{-1/\lambda_{i}}\right)$ where we take the root which has a positive imaginary part.
% Using this result and Eq.~\eqref{eq:smatrix}, we numerically find the Andreev spectrum.

% \comment{The lowest energy at $|k| < k_\textrm{cutoff}$ sets the band gap.}
% We cut off the spectrum at the momentum corresponding to the proposed cutoff in the zigzag, and compare it to the numerically obtained gap for a true zigzag system.
